```{r, echo=FALSE}
options(mmand.display.newDevice=FALSE)
knitr::opts_chunk$set(collapse=TRUE, fig.path="tools/figures/", fig.dim=c(1,1), dpi=128)
```

[![CRAN version](http://www.r-pkg.org/badges/version/loder)](https://cran.r-project.org/package=loder) [![CI](https://github.com/jonclayden/loder/actions/workflows/ci.yaml/badge.svg)](https://github.com/jonclayden/loder/actions/workflows/ci.yaml) [![codecov](https://codecov.io/gh/jonclayden/loder/branch/master/graph/badge.svg)](https://app.codecov.io/gh/jonclayden/loder) [![status](https://tinyverse.netlify.com/badge/loder)](https://cran.r-project.org/package=loder)

# Simple, dependency-free access to PNG images

The `loder` package provides functions for easily reading from PNG image files and writing to them. It functions in a very similar way to Simon Urbanek's venerable [`png` package](https://github.com/s-u/png), but unlike that package it does not require external `libpng` and `zlib` libraries to be installed. Instead, `loder` includes Lode Vandevenne's compact LodePNG library, which provides self-contained PNG read and write functionality.

## Installation

The package may be installed from [CRAN](https://cran.r-project.org/package=loder) using the standard command

```{r, eval=FALSE}
install.packages("loder")
```

Alternatively, the latest development version can be installed directly from GitHub using the `remotes` package, viz.

```{r, eval=FALSE}
## install.packages("remotes")
remotes::install_github("jonclayden/loder")
```

## Reading and writing files

The `readPng` function is used to read PNG images. It returns an array of class `loder` with three dimensions representing rows, columns and channels. Greyscale images have one channel, grey/alpha images have two, red/green/blue (RGB) images have three, and red/green/blue/alpha images have four.

```{r}
library(loder)
path <- system.file("extdata", "pngsuite", "basn6a08.png", package="loder")
image <- readPng(path)
dim(image)
```

Metadata including background colour, aspect ratio and pixel size or image resolution are available through attributes, if they are stored in the file.

```{r}
path <- system.file("extdata", "pngsuite", "cdfn2c08.png", package="loder")
attributes(readPng(path))
```

The `display` function from the [`mmand` package](https://github.com/jonclayden/mmand) respects these attributes, and can be used to view the image on an R graphics device.

```{r}
## install.packages("mmand")
mmand::display(image)
```

Metadata alone may be read using the `inspectPng` function, which also gives information about the compression level achieved:

```{r}
inspectPng(path)
```

Image data can be written back to a file using the `writePng` function. The theoretical numerical range of the pixel intensities (i.e., the black and white points) can be set using the `range` attribute or an argument of the same name.

```{r}
writePng(image, tempfile(), range=c(0,200))
```

This will rescale the image intensities, and clip any values outside the specified range to the appropriate extreme.
